---
title: "edld610_lab3"
author: "Cameron Kay, Lea Frank, Ashley Miller"
date: "1/30/2019"
output:
  html_document:
    highlight: zenburn
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
# set knitr options
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.width = 7.5,
                      fig.height = 4.0)

# load packages
library(rio)
library(here)
library(tidyverse)
library(fivethirtyeight)
library(lubridate)
library(magrittr)
library(ggthemes)
library(ggThemeAssist)
library(waffle)
library(extrafont)
library(viridis)

# disable scientific notation
options(scipen = 999)
```

#### 1) Start with the google trends

+ Visualize the change in trends for each hurricane over time in one plot with three scales. Map color to a discrete scale.

```{r lfrank_part1_plot1}

google <- google_trends %>% 
  gather(key = "hurricane", value = "trends", -1) %>% 
  separate(col = "hurricane", into = c("tmp1","hurricane","tmp2")) %>% 
  select(-tmp1, -tmp2)

google %>% 
  ggplot(aes(x = date, y = trends, color = hurricane)) +
  geom_line() +
  theme_minimal() +
  scale_color_colorblind() +
  labs(title = "Google Search Trends for Hurricanes in 2017",
       y = "Search Frequency",
       x = "Date",
       fill = "Hurricane")
```

+ Visualize the trends again, this time using geom_ribbon, setting the ymin to 0.

```{r lfrank_part1_plot2}
google %>% 
  ggplot(aes(x = date, y = trends, color = hurricane)) +
  geom_ribbon(aes(ymin = 0, ymax = trends, fill = hurricane), alpha = .5) +
  theme_minimal() +
  scale_fill_colorblind() +
  scale_color_colorblind() +
  labs(title = "Google Search Trends for Hurricanes in 2017",
       y = "Search Frequency",
       x = "Date")

# I couldn't figure out how to change the legend title without creating an extra legend for the other scale (color or fill)
```

+ Visualize the same relation, but change the scales so color is mapped to a continuous scale.

```{r lfrank_part1_plot3}
google %>% 
  ggplot(aes(x = date, y = hurricane, color = trends)) +
  geom_point(aes(size = trends)) +
  scale_color_distiller(palette = "Purples", direction = 1) +
  theme_minimal() +
  labs(title = "Google Search Trends for Hurricanes in 2017",
       y = "Hurricanes",
       x = "Date") +
  guides(color = guide_legend(title = "Search Frequency"), size = guide_legend(title = "Search Frequency"))

```

+ Create a data frame that includes the dates that each hurricane made landfall. Annotate one of your first two plots to include these reference points with vertical lines. The hurricanes made landfall on (approximately) the following days:
      Harvey: August 25, 2017
      Irma: September 10, 2017
      Jose: Did not make landfall
      Maria: September 20, 2017
      
```{r lfrank_part1_plot4}
hc_dates <- tibble(hurricane = c("harvey","irma","jose","maria"),
                   landfall = lubridate::ymd(c("2017-08-25","2017-09-10",NA,"2017-09-20")))

google <- left_join(google, hc_dates, by = "hurricane")

google %>% 
  ggplot(aes(x = date, y = trends, color = hurricane, fill = hurricane)) +
  geom_ribbon(aes(ymin = 0, ymax = trends), alpha = .5) +
  geom_vline(aes(xintercept = landfall, color = hurricane), size = .5) +
  theme_minimal() +
  scale_fill_colorblind() +
  scale_color_colorblind() +
  labs(title = "Google Search Trends for Hurricanes in 2017",
       y = "Search Frequency",
       x = "Date")

```
      
+ Make any final modifications to the plot you’d like to communicate the information as effectively as possible

```{r lfrank_part1_plot5}
google$hurricane <- factor(google$hurricane, labels = c("Harvey","Irma","Jose","Maria"))
colnames(google)[2] <- "Hurricane"

fig_cols <- colorblind_pal()(4)

google %>% 
  ggplot(aes(x = date, y = trends, fill = Hurricane)) +
  geom_ribbon(aes(ymin = 0, ymax = trends), alpha = .6) +
  #geom_vline(aes(xintercept = landfall, color = Hurricane), 
  #           size = .3, linetype = "dashed") +
  theme_minimal() +
  scale_fill_colorblind() +
  scale_color_colorblind() +
  labs(title = "Google Search Trends for Hurricanes in 2017",
       subtitle = "Date of landfall indicated by dashed line",
       y = "Search Frequency",
       x = "Date") +
  geom_segment(x = hc_dates$landfall[1], y = 0, xend = hc_dates$landfall[1], yend = 100,
           color = fig_cols[1], size = .3, linetype = "dotted") +
  geom_segment(x = hc_dates$landfall[2], y = 0, xend = hc_dates$landfall[2], yend = 100,
           color = fig_cols[2], size = .3, linetype = "dotted") +
  geom_segment(x = hc_dates$landfall[4], y = 0, xend = hc_dates$landfall[4], yend = 100,
           color = fig_cols[4], size = .3, linetype = "dotted") +
  annotate(geom = "text", y = 105, x = hc_dates$landfall[1], label = "Aug 25") +
  annotate(geom = "text", y = 105, x = hc_dates$landfall[2], label = "Sep 10") +
  annotate(geom = "text", y = 105, x = hc_dates$landfall[4], label = "Sep 20")

```

#### 2) Replicate the “National cable news networks” plot using the tv_states data. Don’t worry about all the labels. Try to match the colors but don’t worry about getting them exact.

```{r part2_amiller, fig.width = 9, fig.height = 5}

#load data
tv_data <- tv_states %>%
  gather(location, share, florida:puerto_rico) 

tv_data$location <- factor(tv_data$location, levels = c("florida", "texas", "puerto_rico"))

dateVec <- as.Date(c("2017-08-15", "2017-08-29", "2017-09-01", "2017-09-09", "2017-09-29", "2017-10-01", "2017-10-10"))

ggplot(tv_data, aes(x = date, y = share, fill = location)) +
  geom_area(color = "white",
            position = "identity") +
  scale_fill_manual(values = c("#E2327D", "#F36A05", "#0BBCB8")) +
  theme_minimal() +
  labs(title = "National cabel news networks",
       x = "", 
       y = "Share of sentences") + 
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 20),
        legend.position = "none",
        axis.title = element_text(size = 11, face = "bold"),
        axis.text = element_text(colour = "gray58")) +
  geom_hline(yintercept = 0) +
  scale_x_date(limits = c(min(dateVec), max(dateVec)), 
               date_labels = "%b. %d") +
  scale_y_continuous(breaks = c(1, 2, 3),
                     label = c("1", "2", "3%"),
                     limits = c(0, NA)) +
  geom_text(y = 0.7, x = dateVec[2], 
         label = "Texas", fontface = 2, size = 5, color = "black") +
  geom_text(y = 1.2, x = dateVec[4], 
           label = "Florida", fontface = 2, size = 5) +
  geom_text(y = 1.2, x = dateVec[5], 
           label = "Puerto Rico", fontface = 2, size = 5)
 
```

#### 3) Use the comic_characters dataset to visualize the proportion of different genders using at least **three** different methods. With your final visualization, make any additional modifications you think would help reduce the cognitive load neccessary to comprehend the figure.

```{r part3_ckay, fig.width = 9, fig.height = 5}
# Use the comic_characters dataset to visualize the proportion of different genders using at least three different methods. With your final visualization, make any additional modifications you think would help reduce the cognitive load neccessary to comprehend the figure.

# load packages



###############################
###VISUALIZATION 1
###############################

# import fonts (run this after you install the font awesome font of the first time)
# font_import()

# load fonts 
loadfonts()

# prepare data
char_mf <- comic_characters %>%
  filter(!is.na(sex)) %>%
  count(sex) %>%
  filter(n > 500) %>%
  mutate(n          = round(n / 1000),
         sex        = gsub(" Characters", "", sex),
         proportion = n / sum(n),
         percent    = round(proportion * 100),
         percent    = paste(as.character(percent), "%", sep = ""))

# visualization 1
sex_count <- char_mf$n

names(sex_count) <- char_mf$sex

names(sex_count)

waffle(parts = sex_count, 
       use_glyph = "shield", 
       rows = 2, 
       glyph_size = 10, 
       title = "Proportion of Comic Book Characters who are female and male",
       xlab = "One shield = 1000 characters",
       colors = c("#8C72F7", "#5BA4D3"),
       legend_pos = "bottom")

# visualization 2
ggplot(char_mf, aes(x = "", y = proportion, fill = sex)) + 
  geom_bar(stat = "identity", width = 1, color = "gray20") + 
  coord_polar("y", start = 0) +
  scale_fill_manual(values = c("#8C72F7", "#5BA4D3")) +
  labs(title = "Proportion of Comic Book Characters who are female and male",
       x     = "",
       y     = "",
       fill  = "Sex") +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
  geom_text(aes(label = percent), position = position_stack(vjust = .5))

#visualization 3
char_mf <- comic_characters %>%
  filter(!is.na(sex)) %>%
  count(sex) %>%
  mutate(sex = gsub(" Characters", "", sex))

ggplot(char_mf, aes(x = fct_reorder(sex, n), y = n, fill = sex)) +
  geom_col() +
  scale_y_log10() +
  scale_fill_viridis_d()
  


```
